The Unicode Cookbook \linebreak for Linguists

### 2018-06-05 book/unicode.tex
work towards paperhivers chapter 8

Note that our two libraries have rather different implementation histories, thus they may not give the same results in all situations (as discussed in Chapter~\ref{orthography-profiles}). However, we do provide extensive test suites for each implementation that follow standard practices to make sure that results are correct. Users should refer to these tests and to the documentation in each release for specifics about each implementation. Note that due to the different naming convention practices in Python and R, function names differ between the two libraries. Also, the performance with larger datasets may not be comparable between the Python and R implementations. In sum, our two libraries should be considered as proof of concept and not as the final word on the practical application of the specifications discussed in the previous chapter. In our experience, the current versions are sufficiently fast and stable to be useful for academic practice (e.g.\ checking data consistency, or analyzing and transliterating small to medium sized data sets), but they should probably not be used for full-scale industry applications without adaptation.

First, in Section \ref{installing-python-and-r} we explain how to install Python\footnote{\url{https://www.python.org/}} and R.\footnote{\url{https://www.r-project.org/}} Then in Sections \ref{python-implementations} and \ref{r-implementation}, we discuss our Python and R software packages, respectively. In addition to the material presented here to get users started, we maintain several case studies online that illustrate how to use orthography profiles in action. For convenience, we make these recipes available as Jupyter Notebooks\footnote{\url{http://jupyter.org/}} in our Github repository.\footnote{\url{https://github.com/unicode-cookbook/}} In the final section in this chapter, we also briefly describe a few recipes that we do not go into detail in this book.

@@

You can also install the \texttt{segments} package with from the Github repository:\footnote{\url{https://github.com/cldf/segments}}

@@

% latex table generated in R 3.5.0 by xtable 1.8-2 package
% Mon Jun  4 10:30:43 2018
\begin{table}[H]
\centering
\begingroup\scriptsize
\begin{tabular}{llll}
  \toprule
Gr. & Freq. & Codepoint & Unicode Name \\ 
  \midrule
Ú & 1 & U+00DA & LATIN CAPITAL LETTER U WITH ACUTE \\ 
  Ú & 1 & U+0055, U+0301 & LATIN CAPITAL LETTER U, COMBINING ACUTE ACCENT \\ 
  Ù & 1 & U+00D9 & LATIN CAPITAL LETTER U WITH GRAVE \\ 
  Ù & 1 & U+0055, U+0300 & LATIN CAPITAL LETTER U, COMBINING GRAVE ACCENT \\ 
  Û & 1 & U+00DB & LATIN CAPITAL LETTER U WITH CIRCUMFLEX \\ 
  Û & 1 & U+0055, U+0302 & LATIN CAPITAL LETTER U, COMBINING CIRCUMFLEX ACCENT \\ 
   \bottomrule
\end{tabular}
\endgroup
\caption{Profile 1 (default settings, splitting grapheme clusters)} 
\label{tab:profile1}
\end{table}

By specifying the splitting separator as the empty string
\texttt{sep~=~""}, it is possible to split the string into Unicode code points,
thus separating the combining diacritics. The idea behind this option
\texttt{sep} is that separating by a character allows for user-determined
separation. The most extreme choice here is the empty string \texttt{sep~=~""},
which is interpreted as separation everywhere. The other extreme is the default
setting \texttt{sep~=~NULL}, which means that the separation is not
user-defined, but relegated to the Unicode grapheme definitions.

@@

% latex table generated in R 3.5.0 by xtable 1.8-2 package
% Mon Jun  4 10:30:43 2018
\begin{table}[H]
\centering
\begingroup\scriptsize
\begin{tabular}{llll}
  \toprule
Grapheme & Frequency & Codepoint & Unicode Name \\ 
  \midrule
́ & 1 & U+0301 & COMBINING ACUTE ACCENT \\ 
  ̀ & 1 & U+0300 & COMBINING GRAVE ACCENT \\ 
  ̂ & 1 & U+0302 & COMBINING CIRCUMFLEX ACCENT \\ 
  U & 3 & U+0055 & LATIN CAPITAL LETTER U \\ 
  Ú & 1 & U+00DA & LATIN CAPITAL LETTER U WITH ACUTE \\ 
  Ù & 1 & U+00D9 & LATIN CAPITAL LETTER U WITH GRAVE \\ 
  Û & 1 & U+00DB & LATIN CAPITAL LETTER U WITH CIRCUMFLEX \\ 
   \bottomrule
\end{tabular}
\endgroup
\caption{Profile 2 (splitting by code points)} 
\label{tab:profile2}
\end{table}

Some characters look identical, although they are encoded differently.
Unicode offers different ways of normalization (see
Section~\ref{pitfall-canonical-equivalence}), which can be invoked here as well
using the option \texttt{normalize}. NFC normalization turns everything into the
precomposed characters, while NFD normalization separates everything into base
characters with combining diacritics. Splitting by code points (i.e. \texttt{sep~=~""}) shows the results of these two normalizations.

\begin{knitrout}\footnotesize
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlcom{# after NFC normalization unicode code points have changed}
\hlstd{profile_3} \hlkwb{<-} \hlkwd{write.profile}\hlstd{(example,} \hlkwc{normalize} \hlstd{=} \hlstr{"NFC"}\hlstd{,} \hlkwc{sep} \hlstd{=} \hlstr{""}\hlstd{)}
\hlcom{# NFD normalization gives yet another structure of the code points}
\hlstd{profile_4} \hlkwb{<-} \hlkwd{write.profile}\hlstd{(example,} \hlkwc{normalize} \hlstd{=} \hlstr{"NFD"}\hlstd{,} \hlkwc{sep} \hlstd{=} \hlstr{""}\hlstd{)}
\end{alltt}
\end{kframe}
\end{knitrout}

% latex table generated in R 3.5.0 by xtable 1.8-2 package
% Mon Jun  4 10:30:43 2018
\begin{table}[H]
\centering
\begingroup\scriptsize
\begin{tabular}{llll}
  \toprule
Grapheme & Frequency & Codepoint & Unicode Name \\ 
  \midrule
Ú & 2 & U+00DA & LATIN CAPITAL LETTER U WITH ACUTE \\ 
  Ù & 2 & U+00D9 & LATIN CAPITAL LETTER U WITH GRAVE \\ 
  Û & 2 & U+00DB & LATIN CAPITAL LETTER U WITH CIRCUMFLEX \\ 
   \bottomrule
\end{tabular}
\endgroup
\caption{Profile 3 (splitting by NFC code points)} 
\label{tab:profile3}
\end{table}

% latex table generated in R 3.5.0 by xtable 1.8-2 package
% Mon Jun  4 10:30:43 2018
\begin{table}[H]
\centering
\begingroup\scriptsize
\begin{tabular}{llll}
  \toprule
Grapheme & Frequency & Codepoint & Unicode Name \\ 
  \midrule
́ & 2 & U+0301 & COMBINING ACUTE ACCENT \\ 
  ̀ & 2 & U+0300 & COMBINING GRAVE ACCENT \\ 
  ̂ & 2 & U+0302 & COMBINING CIRCUMFLEX ACCENT \\ 
  U & 6 & U+0055 & LATIN CAPITAL LETTER U \\ 
   \bottomrule
\end{tabular}
\endgroup
\caption{Profile 4 (splitting by NFD code points)} 
\label{tab:profile4}
\end{table}

It is important to realize that for Unicode grapheme definitions, NFC
and NFD normalization are equivalent. This can be shown by normalizing the
example in either NFC or NFD, but using default separation in
\texttt{write.profile}. To be precise, default separation means setting
\texttt{sep~=~NULL}, but that has not be added explicitly.

\begin{knitrout}\footnotesize
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlcom{# note that NFC and NFD normalization are identical}
\hlcom{# for unicode grapheme definitions}
\hlstd{profile_5} \hlkwb{<-} \hlkwd{write.profile}\hlstd{(example,} \hlkwc{normalize} \hlstd{=} \hlstr{"NFD"}\hlstd{)}
\hlstd{profile_6} \hlkwb{<-} \hlkwd{write.profile}\hlstd{(example,} \hlkwc{normalize} \hlstd{=} \hlstr{"NFC"}\hlstd{)}
\end{alltt}
\end{kframe}
\end{knitrout}

% latex table generated in R 3.5.0 by xtable 1.8-2 package
% Mon Jun  4 10:30:43 2018
\begin{table}[H]
\centering
\begingroup\scriptsize
\begin{tabular}{llll}
  \toprule
Gr. & Freq. & Codepoint & Unicode Name \\ 
  \midrule
Ú & 2 & U+0055, U+0301 & LATIN CAPITAL LETTER U, COMBINING ACUTE ACCENT \\ 
  Ù & 2 & U+0055, U+0300 & LATIN CAPITAL LETTER U, COMBINING GRAVE ACCENT \\ 
  Û & 2 & U+0055, U+0302 & LATIN CAPITAL LETTER U, COMBINING CIRCUMFLEX ACCENT \\ 
   \bottomrule
\end{tabular}
\endgroup
\caption{Profile 5 (splitting by graphemes after NFD)} 
\label{tab:profile5}
\end{table}

% latex table generated in R 3.5.0 by xtable 1.8-2 package
% Mon Jun  4 10:30:43 2018
\begin{table}[H]
\centering
\begingroup\scriptsize
\begin{tabular}{llll}
  \toprule
Gr. & Freq. & Codepoint & Unicode Name \\ 
  \midrule
Ú & 2 & U+00DA & LATIN CAPITAL LETTER U WITH ACUTE \\ 
  Ù & 2 & U+00D9 & LATIN CAPITAL LETTER U WITH GRAVE \\ 
  Û & 2 & U+00DB & LATIN CAPITAL LETTER U WITH CIRCUMFLEX \\ 
   \bottomrule
\end{tabular}
\endgroup
\caption{Profile 6 (splitting by graphemes after NFC)} 
\label{tab:profile6}
\end{table}

Note that these different profiles can also be produced using the bash
executable \texttt{writeprofile} (see above for how to install the bash executable). 
Exactly this example is also included in the help file of the executable.

@@

% latex table generated in R 3.5.0 by xtable 1.8-2 package
% Mon Jun  4 10:30:43 2018
\begin{table}[htb]
\centering
\begingroup\scriptsize
\begin{tabular}{llll}
  \toprule
Grapheme & Frequency & Codepoint & UnicodeName \\ 
  \midrule
sh &  &  &  \\ 
  ch &  &  &  \\ 
  sch &  &  &  \\ 
  ie &  &  &  \\ 
  oo &  &  &  \\ 
  a & 2 & U+0061 & LATIN SMALL LETTER A \\ 
  c & 2 & U+0063 & LATIN SMALL LETTER C \\ 
  e & 3 & U+0065 & LATIN SMALL LETTER E \\ 
  f & 1 & U+0066 & LATIN SMALL LETTER F \\ 
  h & 4 & U+0068 & LATIN SMALL LETTER H \\ 
  i & 3 & U+0069 & LATIN SMALL LETTER I \\ 
  m & 4 & U+006D & LATIN SMALL LETTER M \\ 
  o & 2 & U+006F & LATIN SMALL LETTER O \\ 
  p & 2 & U+0070 & LATIN SMALL LETTER P \\ 
  s & 4 & U+0073 & LATIN SMALL LETTER S \\ 
   \bottomrule
\end{tabular}
\endgroup
\caption{Manually edited profile skeleton} 
\label{tab:profile_skeleton1}
\end{table}

@@

% latex table generated in R 3.5.0 by xtable 1.8-2 package
% Mon Jun  4 10:30:43 2018
\begin{table}[htb]
\centering
\begingroup\scriptsize
\begin{tabular}{lllll}
  \toprule
Left & Grapheme & Frequency & Codepoint & UnicodeName \\ 
  \midrule
mi & s &  &  &  \\ 
   & sh &  &  &  \\ 
   & ch &  &  &  \\ 
   & sch &  &  &  \\ 
   & ie &  &  &  \\ 
   & oo &  &  &  \\ 
   & a & 2 & U+0061 & LATIN SMALL LETTER A \\ 
   & c & 2 & U+0063 & LATIN SMALL LETTER C \\ 
   & e & 3 & U+0065 & LATIN SMALL LETTER E \\ 
   & f & 1 & U+0066 & LATIN SMALL LETTER F \\ 
   & h & 4 & U+0068 & LATIN SMALL LETTER H \\ 
   & i & 3 & U+0069 & LATIN SMALL LETTER I \\ 
   & m & 4 & U+006D & LATIN SMALL LETTER M \\ 
   & o & 2 & U+006F & LATIN SMALL LETTER O \\ 
   & p & 2 & U+0070 & LATIN SMALL LETTER P \\ 
   & s & 4 & U+0073 & LATIN SMALL LETTER S \\ 
   \bottomrule
\end{tabular}
\endgroup
\caption{Orthography profile with contextual specification for <s>} 
\label{tab:profile_skeleton2}
\end{table}

@@

% latex table generated in R 3.5.0 by xtable 1.8-2 package
% Mon Jun  4 10:30:43 2018
\begin{table}[htb]
\centering
\begingroup\scriptsize
\begin{tabular}{lllll}
  \toprule
Left & Grapheme & Right & Class & Replacement \\ 
  \midrule
 & a &  &  & a \\ 
   & c &  &  & c \\ 
   & e &  &  & e \\ 
   & i &  &  & i \\ 
   & n &  &  & n \\ 
   \bottomrule
\end{tabular}
\endgroup
\caption{Orthography profile with empty columns for editing contexts} 
\label{tab:profile_editing_1}
\end{table}

@@

% latex table generated in R 3.5.0 by xtable 1.8-2 package
% Mon Jun  4 10:30:43 2018
\begin{table}[htb]
\centering
\begingroup\scriptsize
\begin{tabular}{lllll}
  \toprule
Left & Grapheme & Right & Class & IPA \\ 
  \midrule
 & c & [ie] &  & tʃ \\ 
   & a &  &  & a \\ 
   & n &  &  & n \\ 
   & c &  &  & k \\ 
   & e &  &  & e \\ 
   & i &  &  & i \\ 
   \bottomrule
\end{tabular}
\endgroup
\caption{Orthography profile with regex as context} 
\label{tab:profile_editing_2}
\end{table}

To use this profile, you have to add the option \texttt{regex = TRUE}. Also note
that we have changed the name of the transliteration-column, so we have to tell
the tokenization process to use this column to transliterate. This is done by
adding the option \texttt{transliterate = "IPA"}.

@@

% latex table generated in R 3.5.0 by xtable 1.8-2 package
% Mon Jun  4 10:30:44 2018
\begin{table}[htb]
\centering
\begingroup\scriptsize
\begin{tabular}{lllll}
  \toprule
Left & Grapheme & Right & Class & IPA \\ 
  \midrule
 & c & Vfront &  & tʃ \\ 
   & a &  &  & a \\ 
   & n &  &  & n \\ 
   & c &  &  & k \\ 
   & e &  & Vfront & e \\ 
   & i &  & Vfront & i \\ 
   \bottomrule
\end{tabular}
\endgroup
\caption{Orthography profile with Class as context} 
\label{tab:profile_editing_3}
\end{table}

>>>>

Note that our two libraries have rather different implementation histories, thus they may not give the same results in all situations (as discussed in Chapter~\ref{orthography-profiles}). However, we do provide extensive test suites for each implementation that follow standard practices to make sure that results are correct. Users should refer to these tests and to the documentation in each release for specifics about each implementation. Note that due to the different naming convention practices in Python and R, function names differ between the two libraries. Also, the performance with larger datasets may not be comparable between the Python and R implementations. In sum, our two libraries should be considered as proofs of concept and not as the final word on the practical application of the specifications discussed in the previous chapter. In our experience, the current versions are sufficiently fast and stable to be useful for academic practice (e.g.\ checking data consistency, or analyzing and transliterating small to medium sized data sets), but they should probably not be used for full-scale industry applications without adaptation.

First, in Section \ref{installing-python-and-r} we explain how to install Python\footnote{\url{https://www.python.org/}} and R.\footnote{\url{https://www.r-project.org/}} Then in Sections \ref{python-implementations} and \ref{r-implementation}, we discuss our Python and R software packages, respectively. In addition to the material presented here to get users started, we maintain several case studies online that illustrate how to use orthography profiles in action. For convenience, we make these recipes available as Jupyter Notebooks\footnote{\url{http://jupyter.org/}} in our GitHub repository.\footnote{\url{https://github.com/unicode-cookbook/}} In the final section in this chapter, we also briefly describe a few recipes that we do not go into detail in this book.

@@

You can also install the \texttt{segments} package from the GitHub repository:\footnote{\url{https://github.com/cldf/segments}}

@@

% latex table generated in R 3.5.0 by xtable 1.8-2 package
% Tue Jun  5 16:38:40 2018
\begin{table}[H]
\centering
\begingroup\scriptsize
\begin{tabular}{llll}
  \toprule
Gr. & Freq. & Codepoint & Unicode Name \\ 
  \midrule
Ú & 1 & U+00DA & LATIN CAPITAL LETTER U WITH ACUTE \\ 
  Ú & 1 & U+0055, U+0301 & LATIN CAPITAL LETTER U, COMBINING ACUTE ACCENT \\ 
  Ù & 1 & U+00D9 & LATIN CAPITAL LETTER U WITH GRAVE \\ 
  Ù & 1 & U+0055, U+0300 & LATIN CAPITAL LETTER U, COMBINING GRAVE ACCENT \\ 
  Û & 1 & U+00DB & LATIN CAPITAL LETTER U WITH CIRCUMFLEX \\ 
  Û & 1 & U+0055, U+0302 & LATIN CAPITAL LETTER U, COMBINING CIRCUMFLEX ACCENT \\ 
   \bottomrule
\end{tabular}
\endgroup
\caption{Profile 1 (default settings, splitting grapheme clusters)} 
\label{tab:profile1}
\end{table}

By specifying the splitting separator as the empty string
\texttt{sep~=~""}, it is possible to split the string into Unicode code points,
thus separating the combining diacritics. The idea behind this option
\texttt{sep} is that separating by a character allows for user-determined
separation. The most extreme choice here is the empty string \texttt{sep~=~""},
which is interpreted as separation everywhere. The other extreme is the default
setting \texttt{sep~=~NULL}, which means that the separation is not
user-defined, but relegated to the Unicode grapheme definitions. The result is 
shown in Table \ref{tab:profile2}.

@@

% latex table generated in R 3.5.0 by xtable 1.8-2 package
% Tue Jun  5 16:38:40 2018
\begin{table}[H]
\centering
\begingroup\scriptsize
\begin{tabular}{llll}
  \toprule
Grapheme & Frequency & Codepoint & Unicode Name \\ 
  \midrule
́ & 1 & U+0301 & COMBINING ACUTE ACCENT \\ 
  ̀ & 1 & U+0300 & COMBINING GRAVE ACCENT \\ 
  ̂ & 1 & U+0302 & COMBINING CIRCUMFLEX ACCENT \\ 
  U & 3 & U+0055 & LATIN CAPITAL LETTER U \\ 
  Ú & 1 & U+00DA & LATIN CAPITAL LETTER U WITH ACUTE \\ 
  Ù & 1 & U+00D9 & LATIN CAPITAL LETTER U WITH GRAVE \\ 
  Û & 1 & U+00DB & LATIN CAPITAL LETTER U WITH CIRCUMFLEX \\ 
   \bottomrule
\end{tabular}
\endgroup
\caption{Profile 2 (splitting by code points)} 
\label{tab:profile2}
\end{table}

Some characters look identical, although they are encoded differently.
Unicode offers different ways of normalization (see
Section~\ref{pitfall-canonical-equivalence}), which can be invoked here as well
using the option \texttt{normalize}. NFC normalization turns everything into the
precomposed characters, while NFD normalization separates everything into base
characters with combining diacritics. Splitting by code points (i.e. \texttt{sep~=~""}) 
shows the results of these two normalizations in Tables \ref{tab:profile3} \& \ref{tab:profile4}.

\begin{knitrout}\footnotesize
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlcom{# after NFC normalization Unicode code points have changed}
\hlstd{profile_3} \hlkwb{<-} \hlkwd{write.profile}\hlstd{(example,} \hlkwc{normalize} \hlstd{=} \hlstr{"NFC"}\hlstd{,} \hlkwc{sep} \hlstd{=} \hlstr{""}\hlstd{)}
\hlcom{# NFD normalization gives another structure of the code points}
\hlstd{profile_4} \hlkwb{<-} \hlkwd{write.profile}\hlstd{(example,} \hlkwc{normalize} \hlstd{=} \hlstr{"NFD"}\hlstd{,} \hlkwc{sep} \hlstd{=} \hlstr{""}\hlstd{)}
\end{alltt}
\end{kframe}
\end{knitrout}

% latex table generated in R 3.5.0 by xtable 1.8-2 package
% Tue Jun  5 16:38:40 2018
\begin{table}[H]
\centering
\begingroup\scriptsize
\begin{tabular}{llll}
  \toprule
Grapheme & Frequency & Codepoint & Unicode Name \\ 
  \midrule
Ú & 2 & U+00DA & LATIN CAPITAL LETTER U WITH ACUTE \\ 
  Ù & 2 & U+00D9 & LATIN CAPITAL LETTER U WITH GRAVE \\ 
  Û & 2 & U+00DB & LATIN CAPITAL LETTER U WITH CIRCUMFLEX \\ 
   \bottomrule
\end{tabular}
\endgroup
\caption{Profile 3 (splitting by NFC code points)} 
\label{tab:profile3}
\end{table}

% latex table generated in R 3.5.0 by xtable 1.8-2 package
% Tue Jun  5 16:38:40 2018
\begin{table}[H]
\centering
\begingroup\scriptsize
\begin{tabular}{llll}
  \toprule
Grapheme & Frequency & Codepoint & Unicode Name \\ 
  \midrule
́ & 2 & U+0301 & COMBINING ACUTE ACCENT \\ 
  ̀ & 2 & U+0300 & COMBINING GRAVE ACCENT \\ 
  ̂ & 2 & U+0302 & COMBINING CIRCUMFLEX ACCENT \\ 
  U & 6 & U+0055 & LATIN CAPITAL LETTER U \\ 
   \bottomrule
\end{tabular}
\endgroup
\caption{Profile 4 (splitting by NFD code points)} 
\label{tab:profile4}
\end{table}

It is important to realize that for Unicode grapheme definitions, NFC
and NFD normalization are equivalent. This can be shown by normalizing the
example in either NFD or NFC, as shown in Tables \ref{tab:profile5} \& \ref{tab:profile6}, 
by using the default separation in
\texttt{write.profile}. To be precise, default separation means setting
\texttt{sep~=~NULL}, but that has not be added explicitly below.

\begin{knitrout}\footnotesize
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlcom{# note that NFC and NFD normalization are identical}
\hlcom{# for Unicode grapheme definitions}
\hlstd{profile_5} \hlkwb{<-} \hlkwd{write.profile}\hlstd{(example,} \hlkwc{normalize} \hlstd{=} \hlstr{"NFD"}\hlstd{)}
\hlstd{profile_6} \hlkwb{<-} \hlkwd{write.profile}\hlstd{(example,} \hlkwc{normalize} \hlstd{=} \hlstr{"NFC"}\hlstd{)}
\end{alltt}
\end{kframe}
\end{knitrout}

% latex table generated in R 3.5.0 by xtable 1.8-2 package
% Tue Jun  5 16:38:40 2018
\begin{table}[H]
\centering
\begingroup\scriptsize
\begin{tabular}{llll}
  \toprule
Gr. & Freq. & Codepoint & Unicode Name \\ 
  \midrule
Ú & 2 & U+0055, U+0301 & LATIN CAPITAL LETTER U, COMBINING ACUTE ACCENT \\ 
  Ù & 2 & U+0055, U+0300 & LATIN CAPITAL LETTER U, COMBINING GRAVE ACCENT \\ 
  Û & 2 & U+0055, U+0302 & LATIN CAPITAL LETTER U, COMBINING CIRCUMFLEX ACCENT \\ 
   \bottomrule
\end{tabular}
\endgroup
\caption{Profile 5 (splitting by graphemes after NFD)} 
\label{tab:profile5}
\end{table}

% latex table generated in R 3.5.0 by xtable 1.8-2 package
% Tue Jun  5 16:38:40 2018
\begin{table}[H]
\centering
\begingroup\scriptsize
\begin{tabular}{llll}
  \toprule
Gr. & Freq. & Codepoint & Unicode Name \\ 
  \midrule
Ú & 2 & U+00DA & LATIN CAPITAL LETTER U WITH ACUTE \\ 
  Ù & 2 & U+00D9 & LATIN CAPITAL LETTER U WITH GRAVE \\ 
  Û & 2 & U+00DB & LATIN CAPITAL LETTER U WITH CIRCUMFLEX \\ 
   \bottomrule
\end{tabular}
\endgroup
\caption{Profile 6 (splitting by graphemes after NFC)} 
\label{tab:profile6}
\end{table}

These different profiles can also be produced using the bash
executable \texttt{writeprofile} (see above for how to install the bash executable). 
This example is also included in the help file of the executable.

@@

% latex table generated in R 3.5.0 by xtable 1.8-2 package
% Tue Jun  5 16:38:40 2018
\begin{table}[htb]
\centering
\begingroup\scriptsize
\begin{tabular}{llll}
  \toprule
Grapheme & Frequency & Codepoint & UnicodeName \\ 
  \midrule
sh &  &  &  \\ 
  ch &  &  &  \\ 
  sch &  &  &  \\ 
  ie &  &  &  \\ 
  oo &  &  &  \\ 
  a & 2 & U+0061 & LATIN SMALL LETTER A \\ 
  c & 2 & U+0063 & LATIN SMALL LETTER C \\ 
  e & 3 & U+0065 & LATIN SMALL LETTER E \\ 
  f & 1 & U+0066 & LATIN SMALL LETTER F \\ 
  h & 4 & U+0068 & LATIN SMALL LETTER H \\ 
  i & 3 & U+0069 & LATIN SMALL LETTER I \\ 
  m & 4 & U+006D & LATIN SMALL LETTER M \\ 
  o & 2 & U+006F & LATIN SMALL LETTER O \\ 
  p & 2 & U+0070 & LATIN SMALL LETTER P \\ 
  s & 4 & U+0073 & LATIN SMALL LETTER S \\ 
   \bottomrule
\end{tabular}
\endgroup
\caption{Manually edited profile skeleton} 
\label{tab:profile_skeleton1}
\end{table}

@@

% latex table generated in R 3.5.0 by xtable 1.8-2 package
% Tue Jun  5 16:38:40 2018
\begin{table}[htb]
\centering
\begingroup\scriptsize
\begin{tabular}{lllll}
  \toprule
Left & Grapheme & Frequency & Codepoint & UnicodeName \\ 
  \midrule
mi & s &  &  &  \\ 
   & sh &  &  &  \\ 
   & ch &  &  &  \\ 
   & sch &  &  &  \\ 
   & ie &  &  &  \\ 
   & oo &  &  &  \\ 
   & a & 2 & U+0061 & LATIN SMALL LETTER A \\ 
   & c & 2 & U+0063 & LATIN SMALL LETTER C \\ 
   & e & 3 & U+0065 & LATIN SMALL LETTER E \\ 
   & f & 1 & U+0066 & LATIN SMALL LETTER F \\ 
   & h & 4 & U+0068 & LATIN SMALL LETTER H \\ 
   & i & 3 & U+0069 & LATIN SMALL LETTER I \\ 
   & m & 4 & U+006D & LATIN SMALL LETTER M \\ 
   & o & 2 & U+006F & LATIN SMALL LETTER O \\ 
   & p & 2 & U+0070 & LATIN SMALL LETTER P \\ 
   & s & 4 & U+0073 & LATIN SMALL LETTER S \\ 
   \bottomrule
\end{tabular}
\endgroup
\caption{Orthography profile with contextual specification for <s>} 
\label{tab:profile_skeleton2}
\end{table}

@@

% latex table generated in R 3.5.0 by xtable 1.8-2 package
% Tue Jun  5 16:38:40 2018
\begin{table}[htb]
\centering
\begingroup\scriptsize
\begin{tabular}{lllll}
  \toprule
Left & Grapheme & Right & Class & Replacement \\ 
  \midrule
 & a &  &  & a \\ 
   & c &  &  & c \\ 
   & e &  &  & e \\ 
   & i &  &  & i \\ 
   & n &  &  & n \\ 
   \bottomrule
\end{tabular}
\endgroup
\caption{Orthography profile with empty columns for editing contexts} 
\label{tab:profile_editing_1}
\end{table}

@@

% latex table generated in R 3.5.0 by xtable 1.8-2 package
% Tue Jun  5 16:38:40 2018
\begin{table}[htb]
\centering
\begingroup\scriptsize
\begin{tabular}{lllll}
  \toprule
Left & Grapheme & Right & Class & IPA \\ 
  \midrule
 & c & [ie] &  & tʃ \\ 
   & a &  &  & a \\ 
   & n &  &  & n \\ 
   & c &  &  & k \\ 
   & e &  &  & e \\ 
   & i &  &  & i \\ 
   \bottomrule
\end{tabular}
\endgroup
\caption{Orthography profile with regex as context} 
\label{tab:profile_editing_2}
\end{table}

To use this profile, you have to add the option \texttt{regex = TRUE}. Also note
that we have changed the name of the transliteration column, so we have to tell
the tokenization process to use this column to transliterate. This is done by
adding the option \texttt{transliterate = "IPA"}.

@@

% latex table generated in R 3.5.0 by xtable 1.8-2 package
% Tue Jun  5 16:38:40 2018
\begin{table}[htb]
\centering
\begingroup\scriptsize
\begin{tabular}{lllll}
  \toprule
Left & Grapheme & Right & Class & IPA \\ 
  \midrule
 & c & Vfront &  & tʃ \\ 
   & a &  &  & a \\ 
   & n &  &  & n \\ 
   & c &  &  & k \\ 
   & e &  & Vfront & e \\ 
   & i &  & Vfront & i \\ 
   \bottomrule
\end{tabular}
\endgroup
\caption{Orthography profile with Class as context} 
\label{tab:profile_editing_3}
\end{table}
